---
title: "Using devRate package on T. solanivora (Lepidoptera: Gelechiidae) in Ecuador"
author: "Francois Rebaudo, using data from Crespo-Perez et al. 2011"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## T. solanivora dataset
Dataset for T. solanivora was retrieved from Crespo-Perez et al.
2011^[Crespo-Pérez, V., Rebaudo, F., Silvain, J.-F. and Dangles, O. (2011)
      Modeling invasive species spread in complex landscapes: the case of
      potato moth in Ecuador. Landscape Ecology, 26, 1447–1461.], using Web
Plot Digitizer^[Rohatgi, A. (2015) WebPlotDigitalizer: HTML5 Based Online Tool
                to Extract Numerical Data from Plot Images.].

    rawDevEggs <- matrix(c(10, 0.031, 10, 0.039, 15, 0.047, 15, 0.059, 15.5, 0.066,
      13, 0.072, 16, 0.083, 16, 0.100, 17, 0.100, 20, 0.100, 20, 0.143, 25, 0.171,
      25, 0.200, 30, 0.200, 30, 0.180, 35, 0.001), ncol = 2, byrow = TRUE)

    rawDevLarva <- matrix(c(10, 0.010, 10, 0.014, 10, 0.019, 13, 0.034, 15, 0.024,
      15.5, 0.029, 15.5, 0.034, 15.5, 0.039, 17, 0.067, 20, 0.050, 25, 0.076,
      25, 0.056, 30, 0.0003, 35, 0.0002), ncol = 2, byrow = TRUE)

    rawDevPupa <- matrix(c(10, 0.001, 10, 0.008, 10, 0.012, 13, 0.044, 15, 0.017,
      15, 0.044, 15.5, 0.039, 16, 0.034, 15.5, 0.037, 16, 0.051, 17, 0.051,
      20, 0.080, 20, 0.092, 25, 0.102, 25, 0.073, 30, 0.005,
      35, 0.0002), ncol = 2, byrow = TRUE)

## Finding models in the literature
Before attempting to fit any model to the empirical data, the devRate function
"devRateFind" search the database for previous articles fitting which models to
the organism, either by Order, Family, or species.

    devRateFind(orderSP = "Lepidoptera")
    devRateFind(familySP = "Gelechiidae")

Then, the "devRateInfo" function provides additional information on these models
and on parameter estimations.

    devRateInfo(eq = taylor_81)

Information from the database can be plotted using the "devRatePlotInfo" function.

    devRatePlotInfo (eq = taylor_81, sortBy = "ordersp",
      ylim = c(0, 0.20), xlim = c(0, 50))

## Fitting models to empirical datasets
The empirical data can be fitted to any model in the database with the
"devRateModel" function and empirical data plotted against the model using the
"devRatePlot" function.

    mEggs <- devRateModel(eq = taylor_81, temp = rawDevEggs[,1], devRate = rawDevEggs[,2],
      startValues = list(Rm = 0.05, Tm = 30, To = 5))
    mLarva <- devRateModel(eq = taylor_81, temp = rawDevLarva[,1], devRate = rawDevLarva[,2],
      startValues = list(Rm = 0.05, Tm = 25, To = 5))
    mPupa <- devRateModel(eq = taylor_81, temp = rawDevPupa[,1], devRate = rawDevPupa[,2],
      startValues = list(Rm = 0.05, Tm = 30, To = 5))

```{r, echo=FALSE, results='asis'}
knitr::kable(data.frame(Stage = c("Eggs", "Larva", "Pupa"), model = rep("Taylor", 3),
  Rm = c(0.1934, 0.06838, 0.0978), Rm_SE = c(0.01781, 0.009473, 0.0087),
  Tm = c(25.3418, 21.40871, 22.0114), Tm_SE = c(0.76013, 0.729058, 0.4556),
  To = c(-6.8939, 5.54497, -4.7515), To_SE = c(0.76853, 0.865261, 0.4715)))
```

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side.

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.
