---
title: "Using devRate package on T. solanivora (Lepidoptera: Gelechiidae) in Ecuador"
author: "Francois Rebaudo, using data from Crespo-Perez et al. 2011"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## T. solanivora dataset
Dataset for T. solanivora was retrieved from Crespo-Perez et al.
2011^[Crespo-Pérez, V., Rebaudo, F., Silvain, J.-F. and Dangles, O. (2011)
      Modeling invasive species spread in complex landscapes: the case of
      potato moth in Ecuador. Landscape Ecology, 26, 1447–1461.], using Web
Plot Digitizer^[Rohatgi, A. (2015) WebPlotDigitalizer: HTML5 Based Online Tool
                to Extract Numerical Data from Plot Images.].

    rawDevEggs <- matrix(c(10, 0.031, 10, 0.039, 15, 0.047, 15, 0.059, 15.5, 0.066,
      13, 0.072, 16, 0.083, 16, 0.100, 17, 0.100, 20, 0.100, 20, 0.143, 25, 0.171,
      25, 0.200, 30, 0.200, 30, 0.180, 35, 0.001), ncol = 2, byrow = TRUE)

    rawDevLarva <- matrix(c(10, 0.010, 10, 0.014, 10, 0.019, 13, 0.034, 15, 0.024,
      15.5, 0.029, 15.5, 0.034, 15.5, 0.039, 17, 0.067, 20, 0.050, 25, 0.076,
      25, 0.056, 30, 0.0003, 35, 0.0002), ncol = 2, byrow = TRUE)

    rawDevPupa <- matrix(c(10, 0.001, 10, 0.008, 10, 0.012, 13, 0.044, 15, 0.017,
      15, 0.044, 15.5, 0.039, 16, 0.034, 15.5, 0.037, 16, 0.051, 17, 0.051,
      20, 0.080, 20, 0.092, 25, 0.102, 25, 0.073, 30, 0.005,
      35, 0.0002), ncol = 2, byrow = TRUE)

## Finding models in the literature


The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output:
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side.

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
