---
title: "Using devRate package on _T. solanivora_ (Lepidoptera: Gelechiidae) in Ecuador"
author: "Francois Rebaudo, using data from Crespo-Perez et al. 2011"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{main}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette exemplifies the use of the devRate package to fit a development rate 
model to an empirical dataset (laboratory experiments at air constant temperatures) 
and to build a simple phenological model.

The preliminary step is to perform an experiment where your arthropod study model is 
reared at different constant air temperatures, and to monitor the time at which 
your individuals change of life stage (e.g., eggs to larva). The monitoring of life 
stages is commonly expressed in development rate, corresponding to the inverse 
of the time to reach the next life stage. The time is often expressed in days. For 
example, if an individual takes 10 days from eggs to larva at 15 degrees Celsius, 
the development rate is 1/10 = 0.1. In this vignette we used a dataset from the 
literature.

The dataset for _T. solanivora_ was retrieved from Crespo-Perez et al.
2011^[Crespo-Pérez, V., Rebaudo, F., Silvain, J.-F. and Dangles, O. (2011)
      Modeling invasive species spread in complex landscapes: the case of
      potato moth in Ecuador. Landscape Ecology, 26, 1447–1461.], using Web
Plot Digitizer^[Rohatgi, A. (2015) WebPlotDigitalizer: HTML5 Based Online Tool
                to Extract Numerical Data from Plot Images.]. The dataset is 
also included in the package in the exTropicalMoth object. This object structure 
may seem complicated as it is composed of the laboratory data and the results of the 
modeling. It was designed to reduce the length of examples in the manual. In this 
vignette we present how to organize your own dataset from scratch. 

## Organizing the dataset

In this example _T. solanivora_ was reared at different temperatures with various 
repetitions for each temperature and different temperatures for the different life 
stages:
- the eggs were reared at 10, 13, 15, 15.5, 16, 17, 20, 25, 30, and 35 degrees
- the larva were reared at 10, 13, 15, 15.5, 17, 20, 25, 30, and 35 degrees
- the pupa were reared at 10, 13, 15, 15.5, 16, 17, 20, 25, 30, and 35 degrees
The development rates were monitored at all these temperatures with various 
repetitions for each temperature.

The resulting dataset may look like the following table (for eggs development), 
where the first column corresponds to the temperature and the second column to 
the development rate.

| Temperature | Development Rate |
|------------:|-----------------:|
|10.0         |0.031             |
|10.0         |0.039             |
|13.0         |0.072             |
|...          |...               |

One possibility is to read directly from your table (e.g., using _read.table()_). 
In this example, we copy-pasted the experimental results to create vectors and 
data frames.

```{r}
### Experimental temperatures and development rate for the eggs
expeTempEggs <- c(10.0, 10.0, 13.0, 15.0, 15.0, 15.5, 16.0, 16.0, 17.0, 20.0, 20.0, 
              25.0, 25.0, 30.0, 30.0, 35.0)
expeDevEggs <- c(0.031, 0.039, 0.072, 0.047, 0.059, 0.066, 0.083, 0.1, 0.1, 0.1, 0.143, 
             0.171, 0.2, 0.2, 0.18, 0.001)
dfDevEggs <- data.frame(expeTempEggs, expeDevEggs)

### Experimental temperatures and development rate for the larva
expeTempLarva <- c(10.0, 10.0, 10.0, 13.0, 15.0, 15.5, 15.5, 15.5, 17.0, 20.0, 25.0, 
                   25.0, 30.0, 35.0)
expeDevLarva <- c(0.01, 0.014, 0.019, 0.034, 0.024, 0.029, 0.034, 0.039, 0.067, 0.05, 
                  0.076, 0.056, 0.0003, 0.0002)
dfDevLarva <- data.frame(expeTempLarva, expeDevLarva)

### Experimental temperatures and development rate for the pupa
expeTempPupa <- c(10.0, 10.0, 10.0, 13.0, 15.0, 15.0, 15.5, 15.5, 16.0, 16.0, 17.0, 
                  20.0, 20.0, 25.0, 25.0, 30.0, 35.0)
expeDevPupa <- c(0.001, 0.008, 0.012, 0.044, 0.017, 0.044, 0.039, 0.037, 0.034, 0.051, 0.051, 0.08, 0.092, 0.102, 0.073, 0.005, 0.0002)
dfDevPupa <- data.frame(expeTempPupa, expeDevPupa)

### Same dataset included in the package in the form of matrices
library("devRate")
data(exTropicalMoth)
str(exTropicalMoth[[1]])
```

## Finding models in the literature

Before attempting to fit any model to the empirical data, the devRate function
"devRateFind" search the database for previous articles fitting which models to
the organism, either by Order, Family, or species. The most used models are
presented at the top of the data.frame.

```{r}
devRateFind(orderSP = "Lepidoptera")
```
    
Here we can see that at the time of thie vignette, campbell_74 model was used 
108 times to characterize the relationship between development rate and temperature 
for the Lepidoptera Order, and the taylor_81 model 22 times.

```{r}
devRateFind(familySP = "Gelechiidae")
```

If we focus on the Family (Gelechiidae), campbell_74 has been used 12 times, and
taylor_81 2 times.

```{r}
devRateFind(species = "Tecia solanivora")
```

Unfortunately, the species _Tecia solanivora_ is not in the package database, so 
that we have to find a closely-related species.

The "devRateInfo" function provides additional information on these models
and on parameter estimations.

```{r}
devRateInfo(eq = taylor_81)
```

Here we can see that _Symmetrischema tangolias_ study by Sporleder et al. 2016 has 
used the taylor_81 model. 

To return only the information on the closely-related species, a specific query 
can be performed on the taylor_81 model.

```{r}
taylor_81$startVal[taylor_81$startVal["genSp"] == "Symmetrischema tangolias",]
```

Information from the database can be plotted using the "devRatePlotInfo" function 
if we want to have a first look on the taylor_81 model.

```{r, fig.width = 6, fig.height = 6}
devRatePlotInfo (eq = taylor_81, sortBy = "ordersp",
  ylim = c(0, 0.20), xlim = c(0, 50))
```

In the case that no models were selected a priori because of the research goal 
(e.g., determination of a thermal metric), the taylor_81 model can be used as 
a candidate model.

## Fitting models to empirical datasets

Now that we have a model candidate and starting parameter estimates from a 
closely-related species, we can start the fitting process with our empirical 
data (NLS fit). The empirical data can be fitted to any model in the database 
with the "devRateModel" function, which returns an object of class "nls".

```{r}
### using the vectors from section "Organizing the dataset"
mEggs01 <- devRateModel(eq = taylor_81, 
  temp = expeTempEggs, 
  devRate = expeDevEggs,
  startValues = list(Rm = 0.05, Tm = 30, To = 5))

mLarva01 <- devRateModel(eq = taylor_81, 
  temp = expeTempLarva, 
  devRate = expeDevLarva,
  startValues = list(Rm = 0.05, Tm = 25, To = 10))
      
mPupa01 <- devRateModel(eq = taylor_81, 
  temp = expeTempPupa, 
  devRate = expeDevPupa,
  startValues = list(Rm = 0.1, Tm = 30, To = 10))

### using the data frames from section "Organizing the dataset"
mEggs02 <- devRateModel(eq = taylor_81, 
  df = dfDevEggs,
  startValues = list(Rm = 0.05, Tm = 30, To = 5))

mLarva02 <- devRateModel(eq = taylor_81, 
  df = dfDevLarva,
  startValues = list(Rm = 0.05, Tm = 25, To = 10))
      
mPupa02 <- devRateModel(eq = taylor_81, 
  df = dfDevPupa,
  startValues = list(Rm = 0.1, Tm = 30, To = 10))

### using the dataset included in the package
mEggs <- devRateModel(eq = taylor_81, 
  temp = exTropicalMoth$raw$eggs[,1], 
  devRate = exTropicalMoth$raw$eggs[,2],
  startValues = list(Rm = 0.05, Tm = 30, To = 5))
      
mLarva <- devRateModel(eq = taylor_81, 
  temp = exTropicalMoth$raw$larva[,1], 
  devRate = exTropicalMoth$raw$larva[,2],
  startValues = list(Rm = 0.05, Tm = 25, To = 10))
      
mPupa <- devRateModel(eq = taylor_81, 
  temp = exTropicalMoth$raw$pupa[,1], 
  devRate = exTropicalMoth$raw$pupa[,2],
  startValues = list(Rm = 0.1, Tm = 30, To = 10))
```
      
The summary of the "devRateModel" can be obtained with the "devRatePrint" function.

```{r}
resultNLS <- devRatePrint(myNLS = mLarva, 
  temp = exTropicalMoth$raw$larva[,1], 
  devRate = exTropicalMoth$raw$larva[,2])
```

Empirical data can be plotted against the model using the "devRatePlot" function.

```{r, fig.width = 6, fig.height = 6}
devRatePlot(eq = taylor_81, 
  nlsDR = mEggs, 
  temp = exTropicalMoth$raw$eggs[,1], 
  devRate = exTropicalMoth$raw$eggs[,2],
  pch = 16, ylim = c(0, 0.2))
      
devRatePlot(eq = taylor_81, 
  nlsDR = mLarva, 
  temp = exTropicalMoth$raw$larva[,1], 
  devRate = exTropicalMoth$raw$larva[,2],
  pch = 16, ylim = c(0, 0.1))
      
devRatePlot(eq = taylor_81, 
  nlsDR = mPupa, 
  temp = exTropicalMoth$raw$pupa[,1], 
  devRate = exTropicalMoth$raw$pupa[,2],
  pch = 16, ylim = c(0, 0.15))
```

## Forecasting phenologies from empirical temperature
From the "nls" object obtained above, we can build a simple phenology model using 
temperature time series (e.g., to forecast the organism outbreaks).
In this example the temperature dataset is built from a Normal distribution of mean
15 and a standard deviation of 1, with a time step of one day. The development
models used are those previously fitted with the Taylor model for the three stages.
We assumed that the average time for female adults to lay eggs was of 1 day. We
simulated 500 individuals, with a stochasticity in development rate centered on
the development rate, with a standard deviation of 0.015 (Normal distribution).

```{r}
forecastTsolanivora <- devRateIBM(
  tempTS = rnorm(n = 100, mean = 15, sd = 1),
  timeStepTS = 1,
  models = list(mEggs, mLarva, mPupa),
  numInd = 50,
  stocha = 0.015,
  timeLayEggs = 1)
print(forecastTsolanivora)
```

Results can be plotted using the "devRateIBMPlot" function. From this simple 
model we can observe that if eggs of the first generation are laid at time t = 0, 
adults should be seen at t = [65:85], if the temperature time series correspond 
to the temperatures experienced by the organism and in the absence of other 
limiting factors.

```{r, fig.width = 6, fig.height = 6}
devRateIBMPlot(ibm = forecastTsolanivora, typeG = "density")
```
