---
title: "Using the devRate package to simulate the phenology of *Helicoverpa armigera*"
author: "François Rebaudo ; IRD UMR EGCE"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sec02}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette exemplifies the use of the devRate package to simulate 
*Helicoverpa armigera* phenology using thermal performance curves available 
in the literature.

```{r, "r001"}
require("devRate")
```

## 1. Models published in the literature

Numerous models have been developed to simulate the phenology of *Helicoverpar armigera*. Most are based on the characterization of development as a function of temperature. Insects are reared at various constant temperatures under laboratory conditions, then a model is fitted to the data to obtain a thermal performance curve. The various publications for which models were retrievable are listed in Table 1.

__Table 1.__ Models for *Helicoverpa armigera*

| Ref                                    | DOI                              | Function name in devRate|
|----------------------------------------|----------------------------------|-------------------------|
| Bartekova and Praslicka, 2006          |10.17221/2768-PPS                 |ha_bartekova2006         |
| Jallow and Matsumura, 2001             |10.1303/aez.2001.427              |ha_jallow2001            |
| Mironidis and Savopoulou-Soultani, 2008|10.1093/ee/37.1.16                |ha_mironidis2008_ls      |
| Mironidis and Savopoulou-Soultani, 2008|10.1093/ee/37.1.16                |ha_mironidis2008_nls     |
| Foley, 1981                            |10.1111/j.1440-6055.1981.tb00993.x|ha_foley1981             |
| Kay, 1981                              |10.1111/j.1440-6055.1981.tb01020.x|ha_kay1981_ls            |
| Kay, 1981                              |10.1111/j.1440-6055.1981.tb01020.x|ha_kay1981_nls           |
| Noor-ul-Ane et al., 2018               |10.1017/S0007485317000724         |ha_noorulane2018_ls      |
| Noor-ul-Ane et al., 2018               |10.1017/S0007485317000724         |ha_noorulane2018_nls     |
| Qureshi et al., 1999                   |10.1303/aez.34.327                |ha_qureshi1999_ls        |

Each model is implemented in this package as a function that returns a list object. Each list is composed of two elements: the first contains the mathematical equation used and the second a list of the parameters of the mathematical equation for each life stage of the insect. The life stages available, the models, and the number of temperatures used for fitting are shown in Table 2.

__Table 2.__ Model specifications for *Helicoverpa armigera*

| Name                | Type      |Temperature (Celsius)                                        | Life stages            |
|---------------------|-----------|-------------------------------------------------------------|------------------------|
| ha_bartekova2006    |Campbell   |20, 25, 30 (3)                                               | eggs, larvae, pupae (3)|
| ha_jallow2001       |Campbell   |10, 13.3, 16.4, 20, 22.5, 25, 27.9, 30.5, 32.5 (9)           | eggs, larvae, pupae (3)|
| ha_mironidis2008_ls |Campbell   |10, 12.5, 15, 17.5, 20, 25, 27.5, 30, 32.5, 35, 37.5, 40 (12)| eggs, larvae, pupae (3)|
| ha_mironidis2008_nls|Lactin-2   |10, 12.5, 15, 17.5, 20, 25, 27.5, 30, 32.5, 35, 37.5, 40 (12)| eggs, larvae, pupae (3)|
| ha_foley1981        |Campbell   |20, 24, 28, 32 (4)                      | post-diapausing and non-diapausing pupae (2)|
| ha_kay1981_ls       |Campbell   |8, 10, 13.3, 17.8, 20.8, 24.4, 27.2, 31.4, 35, 39.4 (10)     | eggs (1)               |
| ha_kay1981_nls      |Davidson   |8, 10, 13.3, 17.8, 20.8, 24.4, 27.2, 31.4, 35, 39.4 (10)     | eggs (1)               |
| ha_noorulane2018_ls |Campbell   |10, 15, 17.5, 20, 25, 27.5, 30, 35, 37.5, 40 (10)            | eggs, larvae, pupae (3)|
| ha_noorulane2018_nls|Briere-2   |10, 15, 17.5, 20, 25, 27.5, 30, 35, 37.5, 40 (10)            | overall (1)            |
| ha_qureshi1999_ls   |Campbell   |15, 20, 25, 30 (4)                                           | eggs, larvae, pupae (3)|

## 2. Simulating phenology

Based on the relationship between development and temperature, and on a temperature time series, it is then possible to run simulations of the pest's phenology.

### 2.1. Temperature time series

Any temperature time series can be used to simulate pest phenology. In the following simulations, the time step is expressed in days. If the temperature time series has a time step of, say, 1h, a time step of 1/24 should be specified. In this example, we'll simulate a theoretical temperature time series using a random generation for the normal distribution with mean equal to 16 and standard deviation equal to 2 (60 days with one temperature data every 6 hours ; 60*4 temperature values).

```{r, "r002", fig.width=7, fig.height=4}
temp <- stats::rnorm(n = 60*4, mean = 16, sd = 2)
par(mar = c(4, 4, 0, 0))
plot(
  x = temp, type = "o", 
  xlab = "Time (6 hours time step)", ylab = "Temperature",
  ylim = c(5, 25)
)
```

### 2.2. Model selection and computation

Now we can run a simulation, selecting our temperature time series and the models of our choice.

```{r, "r003", fig.width=7, fig.height=4}
# Using Bartekova and Praslicka, 2006 model
my_model <- ha_bartekova2006()

# This model was designed for eggs, larvae and pupae, using simple linear
# regression models (following Campbell 1974).
forecastX <- devRateIBMparam(
  tempTS = temp, # the temperature time series described above
  timeStepTS = 1/4, # we  have 4 temperature values per day (every 6 hours)
  eq = list("campbell_74", "campbell_74", "campbell_74"), # the model eq for
    # the three life stages
  myParam = list(
    my_model[["model"]]$pupa, # starting with pupae
    my_model[["model"]]$egg,  # then adults and eggs
    my_model[["model"]]$larva # then larvae and the cycle goes on
  ),
  adultLifeStage = 1, # adult life stage after the first model step (pupae 
    #to adults)
  timeLayEggs = 10, # adults longevity fixed to 10 days
  numInd = 10, # 10 individuals for the simulation
  stocha = 0.015 # a bit of intra-population variability
)
```

### 2.3. Model results and comparison






## Legal information

This work is part of the ACOMPLI project. The ACOMPLI project is part of the Strategic Action Plan for the anticipation of the potential European withdrawal of active substances and the development of alternative crop protection techniques (PARSADA). It is financed by ecological planning funds. The French Ministry of Agriculture cannot be held responsible for the content of this package.






<!-- # devtools::document() -->
<!-- # devtools::check() -->
<!-- # devtools::check_win_devel() -->
<!-- # devtools::build() -->

### TO DO
#### P1- add new models adjusted for H. armigera from the literature // OK
#### P2- modify devRateIBMparam to choose when to insert the timeLayEggs value // OK
#### P3- modify devRateIBMparam to use different models for each life stage // OK with backward compatibility
#### P4- add a vignette to explain the different H. armigera models ans how to use them
