---
title: "Reproducing Shi et al. 2016 results using devRate"
author: "François Rebaudo, using data from Shi et al. 2016"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sec01}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette exemplifies the use of the devRate package i) to fit a development rate 
models to empirical datasets with 10 species using 6 nonlinear models, and ii) to 
compare models based on goodness of fit and the trade-off between the model's 
gooness of fit and its structural complexity.

It reproduces the results of the study by Shi et al. 2016^[Shi, P. J., Reddy, G. 
V., Chen, L., & Ge, F. (2015). Comparison of thermal performance equations in 
describing temperature-dependent developmental rates of insects:(I) empirical 
models. Annals of the Entomological Society of America, 109(2), 211-215.].

## Nonlinear models

| Model name     | Model name in devRate    | 
|----------------|--------------------------|
| Briere-1       |briere1_99                | 
| Briere-2       |briere2_99                |
| Lactin         |lactin2_95                |
| Performance-2  |perf2_11                  |
| Beta           ||
| Ratkowsky      ||

## Datasets

__Table 1.__ Datasets of temperature-dependent development rates used by Shi et al. 2016 
and method used to retrieve the source dataset.

| Sample species               | Source                        | Method            |
|------------------------------|-------------------------------|-------------------|
|_Helicoverpa armigera_        | Wu et al. (2009)              | article not found |
|_Kampimodromus aberrans_      | Broufas et al. (2007)         | Table 3           |
|_Toxorhyynchites brevipalpis_ | Trpis (1972)                  | Table 1           |
|_Bactrocera dorsalis_         | Messenger and Flitters (1958) | Table 1           |
|_Aedes aegypti_               | Gilpin and McClelland (1979)  | article not found |
|_Bemisia tabaci_ (B-biotype)  | Xiang et al. (2007)           | article not found |
|_Lipaphis erysimi_            | Liu and Meng (2000)           | Table 1           |
|_Myzus persicae_              | Liu and Meng (1999)           | Table 1           |
|_Epilachna varivestis_        | Shirai and Yara (2001)        | Table 4           |
|_Drosophila buzzatii_         | de Jong (2010)                | Web Plot Digitizer|

The Wu et al. (2009), Gilpin and McClelland (1979), and Xiang et al. (2007) has not been 
found and were excluded from this analysis.

The Broufas et al. (2007) dataset was retrieved from Table 3 where the mean development 
period is provided in hours and transformed in the inverse of days.

```{r}
broufasDS <- data.frame(
  temp = c(15.0, 17.0, 20.0, 22.0, 25.0, 27.0, 30.0, 33.0, 35.0), 
  devRate = 1/c(595.4, 463.0, 260.5, 222.5, 161.8, 153.1, 136.0, 147.8, 182.1)/24)
```

The Trpis et al. (2007) dataset was retrieved from Table 1 where the mean egg development 
time is provided in hours and transformed in the inverse of days.

```{r}
trpisDS <- data.frame(
  temp = c(14.0, 15.0, 16.0, 17.0, 18.0, 19.0, 20.0, 21.0, 22.0, 23.0, 24.0, 25.0, 26.0, 
           27.0, 28.0, 29.0, 30.0, 31.0, 32.0), 
  devRate = 1/c(209.0, 174.0, 165.0, 125.0, 102.0, 90.0, 76.0, 62.0, 55.0, 50.0, 48.0, 
                44.0, 41.0, 39.0, 38.0, 37.5, 37.0, 38.0, 39.0)/24)
```

The Messenger and Flitters (1958) dataset was retrieved from Table 1 where the temperature 
is in Fahrenheit and transformed into Celsius and the mean egg development time is provided 
in hours and transformed in the inverse of days.

```{r}
messengerDS <- data.frame(
  temp = (c(55.0, 56.0, 57.0, 58.0, 60.0, 62.5, 65.0, 67.5, 70.0, 75.0, 80.0, 85.0, 87.5, 
            90.0, 92.5, 95.0, 96.0, 97.0, 97.5) - 32)/1.8, 
  devRate = 1/c(263.0, 232.0, 170.5, 148.0, 121.3, 95.5, 74.0, 62.5, 51.5, 38.0, 30.5, 
                27.0, 25.0, 24.0, 23.5, 25.0, 26.5, 29.3, 34.3)/24)
```

The Liu and Meng (2000) dataset was retrieved from Table 1 (constant temperatures), and for 
the altae form. 

```{r}
liu1DS <- data.frame(
  temp = c(8.3, 11.3, 14.3, 17.3, 20.1, 22.3, 24.7, 26.5, 28.0, 30.0, 33.0, 35.1), 
  devRate = 1/c(47.5, 26.1, 15.8, 11.2, 8.3, 6.7, 6.2, 5.7, 5.4, 5.1, 5.6, 7.1))
```

The Liu and Meng (1999) dataset was retrieved from Table 1 for the altae form.

```{r}
liu2DS <- data.frame(
  temp = c(6.2, 8.3, 11.3, 14.3, 17.3, 20.1, 22.3, 24.7, 26.5, 28.0, 30.0), 
  devRate = 1/c(50.4, 33.9, 21.5, 14.3, 11.2, 8.0, 6.8, 6.2, 5.8, 6.1, 6.5))
```

The Shirai and Yara (2001) dataset was retrieved from Table 4.

```{r}
shiraiDS <- data.frame(
  temp = c(12.5, 15.0, 17.5, 20.0, 22.5, 25.0, 27.5, 30.0, 32.5), 
  devRate = 1/c(87.0, 55.5, 41.9, 32.4, 26.9, 21.7, 20.1, 20.6, 20.8))
```

The de Jong (2010) dataset was retrieved using Web Plot Digitizer on Figure 1B.

```{r}
deJongDS <- data.frame(
  temp = c(15.1, 16.1, 17.1, 18.0, 21.7, 27.9, 30.1, 31.8), 
  devRate = c(0.0253, 0.0291, 0.0370, 0.0381, 0.0710, 0.1079, 0.1098, 0.0972))
```




<!--  `knitr::kable()`. -->

<!-- ```{r, echo=FALSE, results='asis'} -->
<!-- knitr::kable(head(mtcars, 10)) -->
<!-- ``` -->
